\thetitle

### 2022-04-22 paper/paper.tex
export english
docs

% Language selector
\newif\ifen
\enfalse% Change to false to export Korean

@@

\begin{proof}
\ifen
Optimality follows from the foregoing discussion. Sorting $t$ is $O(m \log m)$. The bottleneck step is the creation of the lookup table for $V[j, h]$ in line \ref{Vcreatedlookuptable}. Each entry is generated in unit time, and the size of the table is $O(Hm)$. 
\else
위 논의에 따라 최적성이 성립한다. $t$를 배열하는 시간은 $O(m \log m)$이다. 목병 단계는 \ref{Vcreatedlookuptable}줄에서 $V[j, h]$의 표를 만드는 것이다. 각 원소를 단위 시간 안에 생성할 수 있으며 표의 크기가 $O(Hm)$이다.
\fi
\end{proof}

@@

\ifen
(Computation time.) The bottleneck step is the creation of the lookup table in line \ref{createdlookuptable}, whose size is $m \times |\mathcal{V}|$. Since
\else
(계산 시간.) \ref{createdlookuptable}줄에서 표를 채우는 것이 목병 단계이며 표의 크기는 $m \times |\mathcal{V}|$이다. 그러면
\fi
\begin{equation}
|\mathcal{V}| = \bar U \times 10^{P} + 1 = \bar U \times 10^ { \bigl\lceil\log_{10}\left(m ^2/ \varepsilon \bar U\right)\bigr\rceil} + 1
\leq\frac{m^2}{\varepsilon} \times \text{\ifen const.\else 상수\fi}
\end{equation}
\ifen
is $O(m^2/ \varepsilon)$, the time complexity is as promised.
\else
는 $O(m^2/ \varepsilon)$이므로 시간 복잡성은 정리와 같다.
\fi%To prevent redundant recursive calls to $G(j, v)$, we need to record its values in a dictionary as they are computed; the space required by the dictionary is $O(|\mathcal{V}| m) = O(m^3 / \varepsilon)$.
\end{proof}

@@

\ifen
The experimental variables in Experiments 1 and 2 were the market size $m$, the choice of algorithm, and (for Algorithm \ref{ellisDP3}) the tolerance $\varepsilon$. For each combination of the experimental variables, we generated \nmarkets~markets, and the optimal portfolio was computed three times, with the fastest of the three repetitions recorded as the computation time. We then report the mean and standard deviation across the \nmarkets~markets. Therefore, each cell of each table below represents a statistic over {\the\numexpr 3 * \nmarkets \relax}~computations. Where applicable, we do not count the time required to sort the entries of $t$. As the simulated-annealing heuristic's computation time is negligible, it was excluded from Experiment 2. 
\else
실험 1과 실험 2의 실험 변수로는 시장의 크기 $m$, 해법 선택, 그리고 (알고리즘 \ref{ellisDP3}인 경우) 허용치 $\varepsilon$을 고려했다. 실험 변수의 각 조합에 대해,  \nmarkets 개의 시장을 생성했으며 최적 포트폴리오를 3번 계산 한 것 중 가장 빠른 것으로 계산 시간으로 기록했다. \nmarkets 개의 시장에 대한 평균과 평균편차를 표에 나타낸다. 따라서 아래 표의 각 칸은 {\the\numexpr 3 * \nmarkets \relax}개의 계산에 대한 통계다. $t$를 배열하는 시간은 고려하지 않았다. 그리고 모의 담금질 휴리스틱의 계산 시간이 무시해도 될 정도이므로 실험 2에서 제외했다. 
\fi

@@

\newcommand{\lastptofcaption}{\ifen
For each value of $m$, \nmarkets~markets were generated, and the computation time was recorded as fastest of three repetitions of the algorithm. The table shows the average time (standard deviation) over the \nmarkets~instances.
\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복하여 그중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (평균편차) 시간이 나타난다.\fi}

@@

\ifen \subsection{Explicit treatment of risk aversion}\else \subsection{위험 회피를 모수화한 모형}\fi
\ifen
The familiar Markowitz portfolio optimization model includes an explicit risk-aversion term, whereas our model has settled for an implicit treatment of risk, namely the tradeoff between selective schools and safety schools inherent in the maximax objective function. However, it is possible to augment the objective function $v(\mathcal{X}) = \operatorname{E}[X]$ to incorporate a variance penalty $\beta \geq 0$ as follows:
\else
익숙한 Markowitz 포트폴리오 최적화 모형에서는 명시적인 위험 회피 항이 등장하며, 본 모형은 maximax 목적함수에 내포된 경쟁적인 학교와 안정 지원 학교 사이의 균형에 의한 위험 관리 요소만 고려했다. 그러나 기본 목적함수인 $v(\mathcal{X}) = \operatorname{E}[X]$에 평균편차 페널티 $\beta \geq 0$를 다음처럼 도입할 수 있다:
\fi
\begin{align}
\begin{split}
v_{\mathrm{\beta}}(\mathcal{X}) &=  \operatorname{E}[X] - \beta \operatorname{Var}(X) \\
&=   \operatorname{E}[X] - \beta \left(  \operatorname{E}[X^2]  -  \operatorname{E}[X]^2 \right) \\
& = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)
 - \beta \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j^2 \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)
  + \beta v(\mathcal{X})^2 \\
  &= v(\mathcal{X}; \tau) +  \beta v(\mathcal{X}; t)^2
\end{split}
\end{align}
\ifen
where $\tau_j = t_j - \beta t_j^2$. Since the first term is itself a portfolio valuation function, and the second is a monotonic transformation of one, we speculate that one of the algorithms given above could be used as a subroutine to trace out the efficient frontier by maximizing $v(\mathcal{X}; t)$ subject to the budget constraint and $v(\mathcal{X}; \tau) \geq \alpha$ for various values of $\alpha \geq 0$. 
\else
단, $\tau_j = t_j - \beta t_j^2$. 이의 첫 번째 항 그 자체가 포트폴리오 가치 함수이며, 두 번째 항은 포트폴리오 가치 함수의 단조 변환이다. 따라서 위에서 제시한 알고리즘을 서브루틴으로 사용하면 효율적 투자선(efficient frontier)을 탐색할 수 있어 보인다. 이때 다양한 $\alpha \geq 0$에 대해 예산 제약식과 $v(\mathcal{X}; \tau) \geq \alpha$ 하에서 $v(\mathcal{X}; t)$을 최대화하는 것이 알고리즘의 요점이다.
\fi

@@

When such a dynamic program is implemented using a lookup table or dictionary memoization, producing the optimal solution requires $O(NC)$-time and -space. Kellerer et al. (2004, \S\,3.3) provide a technique for transforming the dynamic program $Z$ into a divide-and-conquer algorithm that produces the optimal solution in $O(N C)$-time and $O(N + C)$-space, a significant improvement. However, their technique requires the objective function to be additively separable in a certain sense that appears difficult to conform to the college application problem. 
\else
수리 실험 결과에 따르면 동적 계획 해법 성능의 목병 요소는 계산 시간이 아니라 메모리 소모량이다. 이 알고리즘의 메모리 소요를 절감하면 상당히 큰 문제를 풀 수 있다.

>>>>

% Language selector
\newif\ifen
\entrue% Change to false to export Korean

@@

\begin{proof}
\ifen
Optimality follows from the foregoing discussion. Sorting $t$ is $O(m \log m)$. The bottleneck step is the creation of the lookup table for $V[j, h]$ in line \ref{Vcreatedlookuptable}. Each entry is generated in unit time, and the size of the table is $O(Hm)$. 
\else
위 논의에 따라 최적성이 성립한다. $t$를 배열하는 시간은 $O(m \log m)$이다. 병목 단계는 \ref{Vcreatedlookuptable}줄에서 $V[j, h]$의 표를 만드는 것이다. 각 원소를 단위 시간 안에 생성할 수 있으며 표의 크기가 $O(Hm)$이다.
\fi
\end{proof}

@@

\ifen
(Computation time.) The bottleneck step is the creation of the lookup table in line \ref{createdlookuptable}, whose size is $m \times |\mathcal{V}|$. Since
\else
(계산 시간.) \ref{createdlookuptable}줄에서 표를 채우는 것이 병목 단계이며 표의 크기는 $m \times |\mathcal{V}|$이다. 그러면
\fi
\begin{equation}
|\mathcal{V}| = \bar U \times 10^{P} + 1 = \bar U \times 10^ { \bigl\lceil\log_{10}\left(m ^2/ \varepsilon \bar U\right)\bigr\rceil} + 1
\leq\frac{m^2}{\varepsilon} \times \text{\ifen const.\else 상수\fi}
\end{equation}
\ifen
is $O(m^2/ \varepsilon)$, the time complexity is as promised.
\else
는 $O(m^2/ \varepsilon)$이므로 시간 복잡성은 정리와 같다.
\fi%To prevent redundant recursive calls to $G(j, v)$, we need to record its values in a dictionary as they are computed; the space required by the dictionary is $O(|\mathcal{V}| m) = O(m^3 / \varepsilon)$.
\end{proof}

@@

\ifen
The experimental variables in Experiments 1 and 2 were the market size $m$, the choice of algorithm, and (for Algorithm \ref{ellisDP3}) the tolerance $\varepsilon$. For each combination of the experimental variables, we generated \nmarkets~markets, and the optimal portfolio was computed three times, with the fastest of the three repetitions recorded as the computation time. We then report the mean and standard deviation across the \nmarkets~markets. Therefore, each cell of each table below represents a statistic over {\the\numexpr 3 * \nmarkets \relax}~computations. Where applicable, we do not count the time required to sort the entries of $t$. As the simulated-annealing heuristic's computation time is negligible, it was excluded from Experiment 2. 
\else
실험 1과 실험 2의 실험 변수로는 시장의 크기 $m$, 해법 선택, 그리고 (알고리즘 \ref{ellisDP3}인 경우) 허용치 $\varepsilon$을 고려했다. 실험 변수의 각 조합에 대해,  \nmarkets 개의 시장을 생성했으며 최적 포트폴리오를 3번 계산 한 것 중 가장 빠른 것으로 계산 시간으로 기록했다. \nmarkets 개의 시장에 대한 평균과 표준편차를 표에 나타낸다. 따라서 아래 표의 각 칸은 {\the\numexpr 3 * \nmarkets \relax}개의 계산에 대한 통계다. $t$를 배열하는 시간은 고려하지 않았다. 그리고 모의 담금질 휴리스틱의 계산 시간이 무시해도 될 정도이므로 실험 2에서 제외했다. 
\fi

@@

\newcommand{\lastptofcaption}{\ifen
For each value of $m$, \nmarkets~markets were generated, and the computation time was recorded as fastest of three repetitions of the algorithm. The table shows the average time (standard deviation) over the \nmarkets~instances.
\else 각 $m$에 대해 \nmarkets 개의 시장을 생성했으며 알고리즘을 3번 반복하여 그중 최소 계산 시간을 기록했다. 표에서 \nmarkets 개의 인스턴스에 대한 평균 (표준편차) 시간이 나타난다.\fi}

@@

\ifen \subsection{Explicit treatment of risk aversion}\else \subsection{위험 회피를 모수화한 모형}\fi
\ifen
The familiar Markowitz portfolio optimization model includes an explicit risk-aversion term, whereas our model has settled for an implicit treatment of risk, namely the tradeoff between selective schools and safety schools inherent in the maximax objective function. However, it is possible to augment the objective function $v(\mathcal{X}) = \operatorname{E}[X]$ to incorporate a variance penalty $\beta \geq 0$ as follows:
\else
익숙한 Markowitz 포트폴리오 최적화 모형에서는 명시적인 위험 회피 항이 등장하며, 본 모형은 maximax 목적함수에 내포된 경쟁적인 학교와 안정 지원 학교 사이의 균형에 의한 위험 관리 요소만 고려했다. 그러나 기본 목적함수인 $v(\mathcal{X}) = \operatorname{E}[X]$에 표준편차 페널티 $\beta \geq 0$를 다음처럼 도입할 수 있다:
\fi
\begin{align}
\begin{split}
v_{\mathrm{\beta}}(\mathcal{X}) &=  \operatorname{E}[X] - \beta \operatorname{Var}(X) \\
&=   \operatorname{E}[X] - \beta \left(  \operatorname{E}[X^2]  -  \operatorname{E}[X]^2 \right) \\
& = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{￣X}: \\ i > j}} (1 - f_{i}) \Bigr)
 - \beta \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j^2 \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)
  + \beta v(\mathcal{X})^2 \\
  &= v(\mathcal{X}; \tau) +  \beta v(\mathcal{X}; t)^2
\end{split}
\end{align}
\ifen
where $\tau_j = t_j - \beta t_j^2$. Since the first term is itself a portfolio valuation function, and the second is a monotonic transformation of one, we speculate that one of the algorithms given above could be used as a subroutine to trace out the efficient frontier by maximizing $v(\mathcal{X}; t)$ subject to the budget constraint and $v(\mathcal{X}; \tau) \geq \alpha$ for various values of $\alpha \geq 0$. 
\else
단, $\tau_j = t_j - \beta t_j^2$. 이의 첫 번째 항 그 자체가 포트폴리오 가치 함수이며, 두 번째 항은 포트폴리오 가치 함수의 단조 변환이다. 따라서 위에서 제시한 알고리즘을 서브루틴으로 사용하면 효율적 투자선(efficient frontier)을 탐색할 수 있어 보인다. 이때 다양한 $\alpha \geq 0$에 대해 예산 제약식과 $v(\mathcal{X}; \tau) \geq \alpha$ 하에서 $v(\mathcal{X}; t)$을 최대화하는 것이 알고리즘의 요점이다.
\fi

@@

When such a dynamic program is implemented using a lookup table or dictionary memoization, producing the optimal solution requires $O(NC)$-time and -space. Kellerer et al. (2004, \S\,3.3) provide a technique for transforming the dynamic program $Z$ into a divide-and-conquer algorithm that produces the optimal solution in $O(N C)$-time and $O(N + C)$-space, a significant improvement. However, their technique requires the objective function to be additively separable in a certain sense that appears difficult to conform to the college application problem. 
\else
수리 실험 결과에 따르면 동적 계획 해법 성능의 병목 요소는 계산 시간이 아니라 메모리 소모량이다. 이 알고리즘의 메모리 소요를 절감하면 상당히 큰 문제를 풀 수 있다.
